Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-06 response-to-reviewers.tex

% -*- mode: LaTeX; mode: flyspell; -*-
% mode: Tex-Pdf -*-
\documentclass{article}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[bookmarks=TRUE,
            colorlinks,
            pdfpagemode=none,
            pdfstartview=FitH,
            citecolor=black,
            filecolor=black,
            linkcolor=blue,
            urlcolor=black,
            ]{hyperref}
\usepackage[top=15mm, bottom=15mm, left=10mm, textwidth=130mm, marginparwidth=70mm]{geometry}
\usepackage{graphicx}
\usepackage{icomma}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}
\usepackage{xspace}
\usepackage{csquotes}
\usepackage[layout=margin,author=,status=draft]{fixme}

\fxsetup{theme=color}
\newcommand{\skb}[1]{
	\begin{quotation}
		\color{Blue} SKB: #1	
	\end{quotation}
}
\newcommand{\debug}[1]{\marginpar{\small\raggedright\begin{spacing}{1}
      \color{BrickRed} #1
\end{spacing}
}}
\newenvironment{response}
{\slshape}{}

@@

\section{Referee 1}

\subsection{Main comments:}

\begin{enumerate}
\item The paper aims to contribute to an already large
  literature. The recent literature in this field focus on the
  identification of causal effects by using instrumental variable
  approaches (see, e.g., Isphording et al., 2014; Yao and van Ours,
  2015). However, I like the fact that the authors a very clear that
  they are not able to identify causal effects. This is, of course, a
  shortcoming and is limiting the overall contribution of the
  paper. Given the specific focus of International Journal of Manpower
  and the interests of its readership, I find that the paper is
  nevertheless providing a small, but still interesting
  contribution. By analysing data from Estonia and providing results
  on natives’ language skills the paper distinguished itself form
  large parts of the literature and adds some new results to our
  knowledge on the labour market effects of language skills.

  \begin{response}
    No response needed here.
  \end{response}
\item 
  My main concern is with some of their wage results. I find it
  puzzling that they don’t find a wage premium for Russian men who
  have Estonian language skills as they do find a sizeable significant
  effect of about 11 to 17 percent for Russian women and a significant
  positive effect for Estonians who speak Russian. Furthermore, they
  find that Estonian language skills do reduce the probability of
  being unemployed for Russians. Why is there a reduction in the
  probability of being unemployed, but no wage effect? Can we trust
  this result? Looking at the descriptive statistics (see Table 1) the
  wage for Russian men who speak Estonian seems to be higher compared
  to Russian men without the ability to speak Estonian (489.96 vs
  401.59). The authors should either provide more statistics on this
  result (if possible) or a more detailed discussion on potential
  reasons that lead to this finding.

  \begin{response}
    This is a good point.  We have added some more analysis (see
    Section 7), and it appears the story is somewhat more nuanced.
    \fxwarning{ensure it's also in the discussion}
  \end{response}
\item
  \debug{OT: I'll do a rewrite after changes are made, hope it's
    sufficient.} 
  The paper needs to be checked for spelling and grammar
  errors. There are a number of typos and errors, which should be
  corrected in a revision. This would normally be a minor comment, but
  here the authors should take this comment very serious as there are
  so many mistakes.
\end{enumerate}

### 2018-07-06 paper/ijm.tex
response: commented on the RU men 0-premium
introduction improved

%Custom commands
\newcommand{\eur}{\EUR}
\newcommand{\std}[1]{\emph{(#1)}}
\newcommand{\V}{{\ensuremath\checkmark}}
\def\fns{\footnotesize}
\renewcommand*{\vec}[1]{\boldsymbol{#1}}

\fxsetup{theme=color}
\def\tenpc{$^{\ast}$}
\def\fivepc{$^{\ast\ast}$}
\def\onepc{$^{\ast\ast\ast}$}
\newcommand{\legend}{\normalsize{Significance levels:\hspace{1em} \tenpc : 10\% \hspace{1em} \fivepc : 5\% \hspace{1em} \onepc : 1\% \normalsize}}
\def\sep{0.5em}
\def\legendTwo{\multicolumn{3}{l}{\footnotesize{Significance levels
			:\hspace{1em} $\dag$ : 10\% \hspace{1em}
			$\ast$ : 5\% \hspace{1em} $\ast\ast$ : 1\% \normalsize}}}
%SKB: Added for documenting the tables and figures
\newcommand{\modelTwo}{age, age\textsuperscript{2}, education, being married, number of children, inter-ethnic household}
\newcommand{\modelThree}{\modelTwo, industry, occupation ,ownership type of company, number of workers in company, working in the public sector, experience in company}
\newcommand{\modelThreeAdd}{industry, occupation, ownership type of company, number of workers in company, working in the public sector, experience in company}	
\newcommand{\restrictions}{The sample period is from year 2000 to year 2012. Sample is limited to persons
	25-55 year old.}
\newcommand{\agerestrictions}{Sample is limited to persons
	25-55 year old.}
%Modified section numbering	
\renewcommand{\thesection}{\arabic{section}.}
\renewcommand{\thesubsection}{\thesection\arabic{subsection}.}

@@

The role of other languages is
less clear.  It is especially unclear what is the role of
different languages in a multi-lingual economy.  For instance,
\cite{Drinkwater1997} finds
a positive effect for Welsh speaking workers in
Wales but \citet{FrancoisGrin1998}, in contrast,
shows that Italian language skills in Swiss non-Italian speaking
regions are associated with lower wage.  It remains unclear if
this is a true effect or just an omitted variable bias.
It is estimated by \cite{Armstrong2015} that there is a statistically significant premium paid to bilingual people both inside and outside Quebec who speak French as their mother tongue People speaking English as their mother tongue also earn a premium, but it is much smaller outside of Quebec. In some cases it is even statistically not significant.

Few previous studies have provided a comprehensive analysis of a labour
market from the viewpoint of language skills.
\citet{Chiswick1995, chiswick+miller2002, chiswick+miller2007, Bellante1998, Chiswick2010} analyse the effects of dominant language fluency for earnings among immigrants in the United States and
\citet{Chiswick1995}  also analyse it for Canada, Australia and Israel.
\citet{Dustmann2003, shields+price2002} explore of immigrant assimilation into the labour market of the United Kingdom.

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  In case of Estonia,
\citet{Toomet2011} shows that the majority language gives little wage
premium for minority men.  However, he analyses neither women, nor
unemployment.  \citet{Lindemann2013} shows that Estonian fluency
is associated with faster job finding rate while while it does not
explain the lower occupational prestige of Russian speakers.  However,
she only includes recent school graduates in her study.\\

\noindent In this paper, we analyse the relationship between both
majority and minority language skills, and English for both
wage and unemployment.  We analyse Estonia, a linguistically divided
country where a large Russian-speaking community forms almost 30\% of
the population.  We provide long--term trends from the last years
of the planned economy in 1989 till the post-recession time in 2012.

We find that in case of, Estonian language is associated with very
small albeit slowly increasing wage premium over
the years, but only for women.
Russian skills, on the other hand, are clearly related to
higher wage but the relationship is slowly fading. English fluency was
associated with a very large wage premium during the first decade
after the introduction of the market economy, it's importance later
declined, but was still large in 2012.  For
women, the wage premium of each of these languages is roughly equal.
In contrast, the sole language that plays a role in determining
unemployment is Estonian, it's fluency is associated with
approximately 5 percentage points lower probability of being unemployed.
These results suggest that Estonian labour market is
segmented both along gender and the ethnic boundaries, men
more likely occupied in less communication-intense jobs.
This view is
supported by data from a specifically designed ethnic integration
dataset.

>>>>

\fxsetup{theme=color}
%Custom commands
\newcommand{\eur}{\EUR}
\newcommand{\std}[1]{\emph{(#1)}}
\newcommand{\V}{{\ensuremath\checkmark}}
\def\fns{\footnotesize}
\renewcommand*{\vec}[1]{\boldsymbol{#1}}

\def\tenpc{$^{\ast}$}
\def\fivepc{$^{\ast\ast}$}
\def\onepc{$^{\ast\ast\ast}$}
\newcommand{\legend}{\normalsize{Significance levels:\hspace{1em} \tenpc : 10\% \hspace{1em} \fivepc : 5\% \hspace{1em} \onepc : 1\% \normalsize}}
\def\sep{0.5em}
\def\legendTwo{\multicolumn{3}{l}{\footnotesize{Significance levels
			:\hspace{1em} $\dag$ : 10\% \hspace{1em}
			$\ast$ : 5\% \hspace{1em} $\ast\ast$ : 1\% \normalsize}}}
%SKB: Added for documenting the tables and figures
\newcommand{\modelTwo}{age, age\textsuperscript{2}, education, being married, number of children, inter-ethnic household}
\newcommand{\modelThree}{\modelTwo, industry, occupation ,ownership type of company, number of workers in company, working in the public sector, experience in company}
\newcommand{\modelThreeAdd}{industry, occupation, ownership type of company, number of workers in company, working in the public sector, experience in company}	
\newcommand{\restrictions}{The sample period is from year 2000 to year 2012. Sample is limited to persons
	25-55 year old.}
\newcommand{\agerestrictions}{Sample is limited to persons
	25-55 year old.}
%Modified section numbering	
\renewcommand{\thesection}{\arabic{section}.}
\renewcommand{\thesubsection}{\thesection\arabic{subsection}.}

@@

The role of other languages, in particular in multi-lingual economy, is
less clear.  For instance,
\cite{Drinkwater1997} finds
a positive effect for Welsh speaking workers in
Wales and \cite{Armstrong2015} shows that bilingual native French-speakers enjoy
a significant premium both inside and outside of Quebec.
\citet{FrancoisGrin1998}, in contrast,
shows that Italian language skills in Swiss non-Italian speaking
regions are associated with lower wage.  It remains unclear if
this is a true effect or just an omitted variable bias.

Few previous studies have provided a comprehensive analysis of a labour
market from the viewpoint of language skills.
\citet{Chiswick1995, chiswick+miller2002, chiswick+miller2007, Bellante1998, Chiswick2010} analyse the effects of dominant language fluency for earnings among immigrants in the United States
\citet{Chiswick1995}  also do this for Canada, Australia and Israel.
\citet{Dustmann2003, shields+price2002} explore of immigrant assimilation into the labour market of the United Kingdom.

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
In case of Estonia,
\citet{Toomet2011} shows that the majority language gives little wage
premium for minority men.  However, he analyses neither women, nor
unemployment.  \citet{Lindemann2013} shows that Estonian fluency
is associated with faster job finding rate while while it does not
explain the lower occupational prestige of Russian speakers.  However,
she only includes recent school graduates in her study.

In this paper, we analyse the relationship between both
majority and minority language skills, and English for both
wage and unemployment.  We analyse Estonia, a former Soviet republic
and now a rapidly growing and
transforming Easter European country.  It is characterized by a stark linguistic division
where the Russian-speaking community forms almost 30\% of
the population.  We provide long--term trends from the last years
of the planned economy in 1989 till the post-recession time in 2012.

We find that the Estonian language, the majority and the official
language of the country, is associated with very
small albeit slowly increasing wage premium over
the years.  Interestingly, this is true only for women, Russian men enjoy
virtually no premium from the majority language skills.
Russian skills, on the other hand, are clearly related to
higher wage but the relationship is slowly fading.  English fluency was
associated with a very large wage premium early after the country
transitioned to a market economy, later it's importance
declined but remained sizable.

In contrast, the sole language that plays a role in determining
unemployment is Estonian.  It's fluency is associated with
approximately 5 percentage points lower probability of being unemployed.
These results suggest that Estonian labour market is
segmented both along gender and the ethnic boundaries, men
more likely occupied in less communication-intense jobs.
